\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

The objective of this project is to research the implementation of a game development suite in HTML5, similar to Unity or Unreal Engine. HTML5 game development is a growing market, and there are very few game development tools currently supporting the development of them. This project attempts provide a game development suite as a HTML5 based web application, utilising scaling cloud technologies and cloud storage to provide the user with the tools needed to develop, publish and play a HTML5 based game all inside their browser. 

A game engine was created using the Three.js and Bullet Physics libraries, with a strong focus on ease of use for the developers. The Entity-Component-System architectural pattern gives the game engine a very simple structure, making it easy to understand for users new to the development of games.

A rich game editor was developed, influenced heavily by existing editors in Unity and Unreal Engine. This allows the user to build a game in their browser by interfacing with the game engine and storing the game on the cloud. One-click publishing is provided for game developers to quickly publish their game for the public to play.

\chapter*{Declaration}
\addcontentsline{toc}{chapter}{Declaration}
I hereby declare that the work described in this dissertation is, except where otherwise stated, entirely my own work and has not been submitted as an exercise for a degree at this or any other university.

~\\
Signed:

~\\
~\\
\rule{8cm}{0.4pt}

Declan Curran

\today

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
I would like to thank my supervisor Brian Keegan for his guidance, enthusiasm and motivation throughout the extent of the project, and my close friends and family for supporting and encouraging me.
