\chapter{Introduction}
\section{Background}
Game engine suites such as Unity and Unreal Engine are becoming one of the biggest leaders in the game development industry, especially now that they are becoming free for anyone to use.\cite{ue4free} However, an opening in the game development market is currently underutilised; that area is HTML5 based games.

Larger game development companies aren't yet looking at HTML5 for the publishing of their games, but it's only a matter of time until this happens; especially with more applications moving to the cloud every day.\cite{5936687}

The inspiration behind this project is to develop a fully featured game development suite, similar to Unity, running entirely on HTML5. This would not only provide game developers with the ability to publish their games to the HTML5 platform, but also entirely removes the need to download huge software packages such as Unity 5 and Unreal Engine 4, which can be over 14 gigabytes in size.

\section{Project Objectives}
The primary goal of this project is to research and develop a scalable web application that provides hobbyist game developers with a comprehensive and easy to use suite of tools that allow them to create, publish and play games inside their browser. This is achieved by completing each of the following objectives:

\begin{itemize}
	\item Build an efficient and capable game engine with a structure that is easy to understand for beginner developers.
	\item Create a comprehensive and easy to use game editor, providing the tools necessary to build an environment in the game engine.
	\item Provide cloud storage for the games created on the system, with integrated project management tools.
	\item Allow game developers to easily publish their games created on the system.
\end{itemize}

\section{Project Challenges}
The most significant challenges that are involved in the realisation of this project are listed below:
\begin{itemize}
	\item The design and development of an easy to understand game engine that provides 3D graphics and real time physics collisions.
	\item Creating a game editor with a level of functionality and usability similar to Unity or Unreal Engine.
	\item Deploying the application on a scalable cloud platform that is able to handle a large amount of users.
	\item The addition of an efficient cloud storage medium for user created game projects.
	\item The implementation of project management and version control functionality using Git.
\end{itemize}

\section{Document Structure}
This document is structured into the following chapters: