\chapter{Conclusion}

In conclusion, the goal of creating a scalable web application that provides users with the ability to create, publish and play HTML5 based games in their browser was achieved. The games created perform well on a variety of platforms, and feature integrated physics, lighting and shadows provided by the Three.js and Ammo.js libraries.

The game engine and editor lack many advanced features, making it incapable of competing with existing game development tools, however a strong foundation exists. When built upon, this will lead to a powerful game development suite comparable to Unity and Unreal Engine. A list of features planned for the future have been detailed in section \ref{section:futurework}.

The extent of this project and it's research lead to many practical learning opportunities. A deep understanding of the inner workings of the Git version control system was needed to work with both GitPython and the GitHub API, which was useful in the implementation of the Git system into the project. Experience in using the Three.js, Ammo.js and Handlebars.js libraries was obtained, all of which are very powerful libraries that will be useful in the future. Knowledge of the design and implementation of a game engine based on the Entity-Component-System pattern was also acquired.

The most influential knowledge gained while working on this project was the practical experience of using Amazon Web Services to deploy a scalable cloud application with cloud storage. This knowledge will be invaluable to me in the future of my career.
